---
title: "Fizzleworth Hunt Metrics: base vs wing vs cat"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 4.5, dpi = 100)

library(dplyr)
library(ggplot2)
library(kableExtra)

# Consistent variant order / colours everywhere
variant_levels <- c("base", "wing", "cat")
variant_cols   <- c(base = "#4C72B0", wing = "#C44E52", cat = "#55A868")
```

```{r load}
raw <- read.csv("hunts.csv", stringsAsFactors = FALSE, na.strings = c("", "NA", "?"))

hunts <- raw %>%
  mutate(
    variant      = factor(variant, levels = variant_levels),
    start_mind   = suppressWarnings(as.numeric(start_mind)),
    end_mind     = suppressWarnings(as.numeric(end_mind)),
    # mind_gained: use the CSV value, else derive from end - start
    mind_gained  = suppressWarnings(as.numeric(mind_gained)),
    mind_gained  = ifelse(is.na(mind_gained) & !is.na(end_mind) & !is.na(start_mind),
                          end_mind - start_mind, mind_gained),
    # recompute the rate metrics so they stay correct as rows are appended
    exp_per_min      = ifelse(!is.na(mind_gained) & duration_min > 0,
                              mind_gained / duration_min, NA_real_),
    kills_per_min    = ifelse(duration_min > 0, kills / duration_min, NA_real_),
    kills_per_100mind = ifelse(!is.na(mind_gained) & mind_gained > 0,
                               kills / mind_gained * 100, NA_real_),
    dmg_per_mana     = ifelse(mana_out > 0, dmg_dealt / mana_out, NA_real_),
    # incoming-danger signal (see README: maneuvers are the real threat, not dmg)
    danger_events    = coalesce(knockdowns, 0) + coalesce(webbed_by_enemy, 0) +
                       coalesce(wounds_taken, 0) + coalesce(stun_events, 0),
    # mana_in minus Symbol-of-Mana refills (flat 50 each). For base/wing this is
    # the sac4mana "sacrifice mana" surge; for cat it also includes Dark
    # Catalyst's 1-14 refund per cast, so read cat's value as an upper bound.
    mana_recovered   = coalesce(mana_in, 0) - 50 * coalesce(symbol_mana_casts, 0),
    got_sac_harvest  = variant != "cat" & mana_recovered > 0,
    hunt_id          = row_number()
  )

n_by_variant <- hunts %>% count(variant, name = "hunts", .drop = FALSE)
```

This report compares hunt performance across the three attack scripts. It reads
`hunts.csv` directly, so re-knitting after you append rows updates every table
and figure. Rate metrics (`exp_per_min`, `kills_per_min`, …) are recomputed here
from the raw columns rather than trusting the CSV's cached values.

**Variants** (see `README.md` for the full rotations):

- **base** — Web → Energy Maelstrom → Tenebrous Tether. No wings.
- **wing** — base rotation + the Energy Wings kit (AOE / reactive Prismatic Aegis).
- **cat** — Corrupt Essence → Dark Catalyst insta-kill on casters; `;reanim` +
  Pain on inciters. Completely different approach. Experimental.

## Sample sizes

```{r counts}
n_by_variant %>%
  kbl(caption = "Hunts logged per variant") %>%
  kable_styling(full_width = FALSE)
```

## All hunts

```{r all-hunts}
hunts %>%
  transmute(
    id = hunt_id, date, variant,
    dur_min = round(duration_min, 2),
    kills,
    mind_gained,
    exp_per_min       = round(exp_per_min, 1),
    kills_per_min     = round(kills_per_min, 2),
    kills_per_100mind = round(kills_per_100mind, 2),
    dmg_per_mana      = round(dmg_per_mana, 2),
    deaths,
    danger_events,
    mana_recovered,
    cube = martial_prowess
  ) %>%
  kbl(caption = "One row per hunt (NA = not measurable from that log)") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))
```

## Summary by variant

```{r summary-fn}
summarise_metric <- function(df, col) {
  fix <- function(x) ifelse(is.nan(x) | is.infinite(x), NA_real_, round(x, 2))
  df %>%
    group_by(variant, .drop = FALSE) %>%
    summarise(
      n_obs  = sum(!is.na(.data[[col]])),
      mean   = fix(mean(.data[[col]], na.rm = TRUE)),
      median = fix(median(.data[[col]], na.rm = TRUE)),
      min    = fix(min(.data[[col]], na.rm = TRUE)),
      max    = fix(max(.data[[col]], na.rm = TRUE)),
      .groups = "drop"
    )
}

metrics <- c("exp_per_min", "kills_per_min", "kills_per_100mind",
             "dmg_per_mana", "duration_min", "danger_events", "mana_recovered")
```

```{r summary-tables, results="asis"}
for (m in metrics) {
  cat(
    summarise_metric(hunts, m) %>%
      kbl(caption = m) %>%
      kable_styling(full_width = FALSE,
                    bootstrap_options = c("striped", "condensed")),
    "\n"
  )
}
```

## Primary metric: exp_per_min

`exp_per_min = mind_gained / duration_min` — field exp earned per minute of
hunt, i.e. leveling speed. Only hunts where both start and end mind were
readable from the log contribute a point.

```{r exp-plot}
plot_metric <- function(df, col, ylab) {
  d <- df %>% filter(!is.na(.data[[col]]))
  ggplot(d, aes(variant, .data[[col]], colour = variant)) +
    geom_jitter(width = 0.12, height = 0, size = 2.6, alpha = 0.85) +
    stat_summary(fun = mean, geom = "crossbar", width = 0.45,
                 linewidth = 0.4, colour = "grey30") +
    scale_colour_manual(values = variant_cols, drop = FALSE) +
    labs(x = NULL, y = ylab,
         subtitle = "points = individual hunts; bar = mean") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
}

plot_metric(hunts, "exp_per_min", "field exp / min")
```

## Kill rate and mana efficiency

```{r other-plots, fig.height=4}
plot_metric(hunts, "kills_per_min", "kills / min")
plot_metric(hunts, "kills_per_100mind", "kills per 100 mind spent")
plot_metric(hunts, "dmg_per_mana", "damage dealt / mana spent")
```

## Incoming danger

Per the README, Sorcerers defend SMR maneuvers poorly and bolts/CS well, so
knockdowns + enemy webs + wounds + stuns are the real danger signal — not
`dmg_taken`. Deaths are shown alongside.

```{r danger}
hunts %>%
  group_by(variant, .drop = FALSE) %>%
  summarise(
    hunts          = n(),
    deaths         = sum(coalesce(deaths, 0L)),
    knockdowns     = sum(coalesce(knockdowns, 0L)),
    enemy_webs     = sum(coalesce(webbed_by_enemy, 0L)),
    wounds         = sum(coalesce(wounds_taken, 0L)),
    stuns          = sum(coalesce(stun_events, 0L)),
    danger_per_hunt = round(mean(danger_events), 2),
    .groups = "drop"
  ) %>%
  kbl(caption = "Incoming danger, totals per variant") %>%
  kable_styling(full_width = FALSE)
```

```{r danger-plot, fig.height=4}
plot_metric(hunts, "danger_events", "danger events (knockdown + web + wound + stun)")
```

## sac4mana harvest

The wing verbs keep the attack script continuously busy, so sac4mana's
"not mid-fight" guard rarely opens *during* combat — a harvest only lands if the
hunt ends with mana still low and a quiet tail (same as a base "tail harvest").
`mana_recovered` = `mana_in` − Symbol-of-Mana refills (50 each); for `cat` it
also folds in Dark Catalyst's 1–14 refund per cast, so treat cat's number as an
upper bound and lean on the `notes` column for whether the real `sacrifice mana`
actually fired.

```{r harvest}
hunts %>%
  group_by(variant, .drop = FALSE) %>%
  summarise(
    hunts                = n(),
    hunts_with_harvest   = if (first(variant) == "cat") NA_integer_ else sum(got_sac_harvest),
    total_mana_recovered = sum(mana_recovered),
    .groups = "drop"
  ) %>%
  kbl(caption = "sac4mana harvest (base/wing); cat = see notes") %>%
  kable_styling(full_width = FALSE)
```

## Confounds to keep in mind

- **Moonstone cube** (`martial_prowess`): hunt 1 ran without it, every hunt
  after with it. Keep it constant across whatever you're actually comparing.
- **Rested vs partial start**: several hunts started at ~10% mind, giving short
  durations and inflated `kills_per_min`. `kills_per_100mind` and `exp_per_min`
  normalise for this; raw `kills` / `duration` do not.
- **Small n**: with ~3 usable `exp_per_min` points per variant and large
  hunt-to-hunt variance, treat the summary means as directional, not
  conclusive.
- `dmg_dealt` is inflated when mobs submerge and get re-hit; treat
  `dmg_per_mana` as directional.
